Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-05-03 WholePlotTrajectories.Rmd
discussion

Both richness and evenness after disturbance followed asymptotic trajectories, sharply increasing for 15 years before stabilizing at higher values than those of control plots.
An increase of communities evenness stems from a higher homogeneity of species distribution.
After disturbance, communities are made of the old, pre-disturbance survivors and the newly recruited trees: changes in composition and abundance distribution are to be found in the recruitment processes and in the pre-disturbance survivors mortality.
The composition of old survivors proved to mirror the initial community [@Herault2018], which leads to support a species turnover among recruited trees with enhanced growth and survival of previously infrequent species.
Indeed, the increase in taxonomic diversity was accompanied by an increase of taxonomic dissimilarity with plots initial state and a functional shift towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998;@Wright2004;@Reich2014].
Disturbance then causes a reorganization of the typical high dominance structure of hyperdiverse mature forests after disturbance, benefiting to pioneers and light demanding species. 
Likely, the changes in abiotic environment and competitive pressure favored pioneers which outcompete others in non limiting resources but are excluded in mature forests by long-lived, more resistant and shade tolerant species.

As stated by the IDH, communities dynamics after disturbance then rely on species functional strategy and corresponding ability to fill the environmental niches made available by disturbance.
Recruited species then mix with pre-disturbance ones from which they differ and constitue a community all the more diversified that the disturbance is intense [@Molino2001].

@@

The trajectories of almost all leaf and stem functional traits, life history traits, and functional diversity followed hump-back trajectories whose maximum were positively correlated to the disturbance intensity, and were recovering towards initial values after 30 years. 
Functional traits and communities functional diversity are the most direct link between biodiversity and ecosystem functioning [@Diaz2005]: these trajectories reflected a consistent recovery in the long term of ecosystem processes [@Guariguata2001].

Same hump-back trajectories were followed by the composition of plots species and functoinal composition.
This similarly meant a recovery of the pre-disturbance state and the maintenance of the initial differences among plots which supports the idea of a convergence of communities composition [@Hubbell1999;@Molino2001;@Baraloto2012a].
Although plots followed homogeneous trends these initial differences in compositon difference, and probably also some abiotic parameters, strongly determined the maximum and time path of the trajectories [@Anderson2007]•.

@@

Despite the consistent recovery of initial diversity and composition there was a time lag between the trajectories followed by taxonomic and functional characteristics.
Taxonomic composition had longer time path and while communities had recovered their functional characteristics, their taxonomic composition and evenness remained altered by disturbance.

This delay between functional and taxonomic dynamics was already observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
The functional diversity of communities rely on the functional strategy of dominant species, according to the "vegetation quantity effect" [@Grime1998].
The functional trajectories of communities are then driven first by the increase of diversity and evenness of dominant species and then by the recruitment of species resembling the old pre-disturbance community which reduced the overall diversity.
Although dominant species restored the functional pre-disturbance characteristics, communities composition remained altered and the infrequent species of initial communities remain missing to achieve plots recovery.
According to the functional traits trajectories, those infrequent species would match the dominant functional type and would participate to the functional redundancy.
The functional redundancy, thas is the functional overlap between species that is typical of the huge biodiversity of tropical forests [@Bellwood2006], therefore was not restored 30 years after disturbance and this is major to consider as it defines forests' resilience [@Trenbath1999;@Elmqvist2003;@Diaz2005].

@@

Our study showed the significant impact of disturbance on tropical forests communities.
It validated the consistency of the IDH in the long term and revealed the contrasting response of taxonomic and functional characteristics with the decoupling between communities taxonomic evenness and their functional diversity and th dominant functional strategies.
Communities recovery remained unachieved but consistent recovery of communities assembly for the lowest disturbance intensity but questioned it after higher disturbance.

>>>>

Both richness and evenness followed asymptotic trajectories after disturbance, sharply increasing for 15 years before stabilizing at higher values than those of control plots.
An increase of communities evenness stems from a higher homogeneity of species distribution.
After disturbance, communities are made of the old, pre-disturbance survivors and the newly recruited trees: changes in composition and abundance distribution are to be found in the recruitment processes and in the pre-disturbance survivors mortality.
Because the composition of old survivors proved to mirror the initial community [@Herault2018], a specific turnover was expected among recruited trees with enhanced growth and survival of previously infrequent species.
Indeed, the increase in taxonomic diversity was accompanied by an increase of taxonomic dissimilarity with plots initial state and a functional shift towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998;@Wright2004;@Reich2014].
Disturbance then causes a reorganization of the typical high dominance structure of hyperdiverse mature forests after disturbance, benefiting to pioneers and light demanding species. 
Likely, the changes in abiotic environment and competitive pressure favored pioneers which outcompete others in non limiting resources but are excluded in mature forests by long-lived, more resistant and shade tolerant species.

As stated by the IDH, communities dynamics after disturbance relied on species functional strategy and corresponding ability to fill the environmental niches made available by disturbance.
Recruited species then mixed with pre-disturbance ones, from which they differed, and constituted a community all the more diversified that the disturbance was intense [@Molino2001].

@@

All diversity and composition trajectories followed similar hump-back trajectories returning towards their initial values.
Communities trajectories therefore illustrated their taxonomic and function resilience.

Because functional traits and functional diversity are the most direct link between biodiversity and ecosystem functioning [@Diaz2005], their resilience meant a consistent recovery of ecosystem processes in the long term [@Guariguata2001].
The resilience of the taxonomic diversity and composition similarly translated  the recovery of communities such as before disturbance.
This resilience meant the maintenance of communities initial differences in composition and supported the idea of their convergence towards a determined state [@Hubbell1999;@Molino2001;@Baraloto2012a].
If all plots followed similar hump-back trajectories, though, some significant differences emerged.
To name but one, the taxonomic evenness of one plot displayed an extensive increase until 30 years after disturbance.
The differences trajectories maxima and time path, along with the fact intrinsinc communities differences are maintained, suggested that the trajectories depend on the initial plots composition or on some abiotic parameters [@Anderson2007].

@@

Despite the recovery of pre-disturbance state there was a time lag between the trajectories followed by communities taxonomic and functional characteristics.
The trajectory of taxonomic composition and evenness had a longer time-path and remained altered while communities functional composition and diversity had already recovered.

Such delay between functional and taxonomic dynamics was already observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015;@Guariguata2001].
According to the "vegetation quantity effect" [@Grime1998], the functional diversity of communities rely on the functional strategy of dominant species.
Communities functional trajectories were then driven first by the increase of the diversity and evenness of dominant species restoring the pre-disturbance functional characteristics.
Then, the functional trajectories were driven by the recruitment of species resembling the old pre-disturbance community, which reduced the overall diversity.
At that time, the taxonomic composition and diversity remained altered: the initially infrequent, functionally redundant species, therefore remained missing.
The functional redundancy, the functional overlap between species that is typical of the huge biodiversity of tropical forests [@Bellwood2006], was then not restored 30 years after disturbance and this is major to consider as it defines forests' resilience [@Trenbath1999;@Elmqvist2003;@Diaz2005].

@@

Our study showed the significant impact of disturbance on tropical forests communities and validated the consistency of the IDH in the long term.
It revealed the contrasting response of communities taxonomic and functional characteristics, with persisting impacts on the species abundance distibution while the functional diversity and dominant functional strategies were restored.
Communities recovery therefore remained unachieved but consistent for the range of disturbance studied here.
The length of the recovery, however, severly questioned the sustainability of intense selective logging and advocated felling cycle much longer than 30 years [@Gourlet-Fleury2005].
